---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Unified Machine Learning Language (mlLang)

<!--
[![Build Status](https://travis-ci.org/sfeuerriegel/SentimentAnalysis.svg?branch=master)](https://travis-ci.org/sfeuerriegel/mlLang)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/mlLang)](https://cran.r-project.org/package=SentimentAnalysis)
[![Coverage Status](https://img.shields.io/codecov/c/github/sfeuerriegel/mlLang/master.svg)](https://codecov.io/github/sfeuerriegel/mlLang?branch=master)
-->

**mlLang** takes hold of the Unified Machine Learning Language (mlLang) inside R. 

TODO: motivating text why we need this

## Overview

The most important functions in **mlLang** are:

* `parseMlTask(path)` converts XML files written in the Unified Machine Learning Language into R objects. Afterwards, `executeMlTask(model, data)` runs the corresponding training routines on the given data.

* Additional function replace the training process for machine learning models in order to convert it into XML files following the Unified Machine Learning Language. 

* This package supports the main functionality as provided by the package ``caret`` for machine learning in R. 

To see examples of these functions in use, check out the help pages, the demos and the vignette. 


## Installation

Using the **devtools** package, you can easily install the latest development version of **mlLang** with

```{r,eval=FALSE}
install.packages("devtools")

# Option 1: download and install latest version from ‘GitHub’
# Not yet supported: devtools::install_github("sfeuerriegel/mlLang")

# Option 2: install directly from bundled archive
devtoos::install_local("mlLang_0.0.1.tar.gz")
```

Note: A CRAN version has not yet been released.

## Usage

This section shows the basic functionality of how to work with the Unified Machine Learning Language. First, load the corresponding package **mlLang**. 

```{r, message=FALSE}
library(mlLang)
```

## Parser: XML to R

The following code demonstrates some of the functionality provided by **mlLang**. 

```{r}
# sample data
data(iris)

# specify sample file
file <- system.file("XML", "iris_classification.xml", package = "mlLang")

# read task from unified machine learning language
task <- parseMlTask(file)

# train object with data
model <- executeMlTask(task, iris)

summary(model)
```

### Simultaneous execution of multiple tasks

```{r, warning=FALSE}
library(tidyverse)

set.seed(0)
in_train <- createDataPartition(iris$Species, p = 0.8, list = FALSE)

files <- list.files(system.file("XML", package = "mlLang"), 
                    pattern = "iris_classification(.*).xml",
                    full.names = TRUE)
models <- runMlTasks(files, iris[in_train, ], logger = NULL)

# best model
perf <- unlist(lapply(models, 
                      function(m) {
                        pred <- predict(m, newdata = iris[-in_train, ] %>% dplyr::select(-Species))
                        cm <- confusionMatrix(pred, iris[-in_train, "Species"])
                        return(cm$overall[["Accuracy"]])
                      }))
basename(files[which.max(perf)])
```

### Batch execution

```{r}
library(BatchJobs)

reg <- makeRegistry(id = "mlLang")
ids <- scheduleMlTasks(reg, files, iris)

showStatus(reg)
```

```{r eval=FALSE}
# retrieve result for first model once completed
m1 <- loadResult(reg, 1)
```


## Generator: R to XML

```{r}
set.seed(0)

data(iris)
predictors <- c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")

task <- newMlTask()

fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 3)

model <- task$train(x = iris[, predictors], y = iris$Species,
                    y_name = "Species",
                    method = "gbm",
                    preProcess = c("center", "scale"),
                    trControl = fitControl,
                    verbose = FALSE)

summary(model)

write(task, "iris_classification.xml")
unlink("iris_classification.xml")
```


## XML syntax of the Unified Machine Learning Language


```{r}
file_correct <- system.file("XML", "iris_classification.xml", package = "mlLang")
checkSchema(file_correct) # should work correctly

file_error <- system.file("XML", "iris_error.xml", package = "mlLang")
checkSchema(file_error) # should give a warning
```


## License

**mlLang** is released under the [MIT License](https://opensource.org/licenses/MIT)

Copyright (c) 2016 Andreas Frorath & Stefan Feuerriegel
