---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Unified Machine Learning Language (mlLang)

<!--
[![Build Status](https://travis-ci.org/sfeuerriegel/SentimentAnalysis.svg?branch=master)](https://travis-ci.org/sfeuerriegel/mlLang)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/mlLang)](https://cran.r-project.org/package=SentimentAnalysis)
[![Coverage Status](https://img.shields.io/codecov/c/github/sfeuerriegel/mlLang/master.svg)](https://codecov.io/github/sfeuerriegel/mlLang?branch=master)
-->

**mlLang** takes hold of the Unified Machine Learning Language (mlLang) inside R. We have developed **mlLang** as an XML-based, unified language for machine learning. It standardizes all relevant steps to train superior models: preprocessing operations, model specification, and the tuning process. It thereby makes model tuning reproducible and documents the underlying process.

This package ships the converter for R. For this purpose, it implements converters in two directions. (1) It automatically reads files in the unified machine learning language from custom XML files and then constructs a corresponding machine learning model in R. (2) It also supports the other direction an automatically converts machine learning models into XML files according to the unified machine learning language. All  machine learning models are built on top of "caret".

Simply load **mlLang** when starting your programming session. Afterwards, all machine learning operations are recorded and written to the disk in an open XML format. This file can be later loaded to reproduce models and training processes from machine learning.

## Overview

The most important functions in **mlLang** are:

* `parseMlTask(path)` converts XML files written in the Unified Machine Learning Language into R objects. Afterwards, `executeMlTask(model, data)` runs the corresponding training routines on the given data.

* Additional function replace the training process for machine learning models in order to convert it into XML files following the Unified Machine Learning Language. 

* This package supports the main functionality as provided by the package ``caret`` for machine learning in R. 

To see examples of these functions in use, check out the help pages, the demos and the vignette. 


## Installation

Using the **devtools** package, you can easily install the latest development version of **mlLang** with

```{r,eval=FALSE}
install.packages("devtools")

# Option 1: download and install latest version from ‘GitHub’
# Not yet supported: devtools::install_github("sfeuerriegel/mlLang")

# Option 2: install directly from bundled archive
devtoos::install_local("mlLang_0.0.1.tar.gz")
```

Note: A CRAN version has not yet been released.

## Usage

This section shows the basic functionality of how to work with the Unified Machine Learning Language. First, load the corresponding package **mlLang**. 

```{r, message=FALSE}
library(mlLang)
```

## Parser: XML to R

The following code demonstrates some of the functionality provided by **mlLang**. 

```{r}
# sample data
data(iris)

# specify sample file
file <- system.file("XML", "iris_classification.xml", package = "mlLang")

# read task from unified machine learning language
task <- parseMlTask(file)

# train object with data
model <- executeMlTask(task, iris)

summary(model)
```

### Simultaneous execution of multiple tasks

```{r, warning=FALSE}
library(tidyverse)

set.seed(0)
in_train <- createDataPartition(iris$Species, p = 0.8, list = FALSE)

files <- list.files(system.file("XML", package = "mlLang"), 
                    pattern = "iris_classification(.*).xml",
                    full.names = TRUE)
models <- runMlTasks(files, iris[in_train, ], logger = NULL)

# best model
perf <- unlist(lapply(models, 
                      function(m) {
                        pred <- predict(m, newdata = iris[-in_train, ] %>% dplyr::select(-Species))
                        cm <- confusionMatrix(pred, iris[-in_train, "Species"])
                        return(cm$overall[["Accuracy"]])
                      }))
basename(files[which.max(perf)])
```

### Batch execution

```{r eval=FALSE}
library(BatchJobs)

reg <- makeRegistry(id = "mlLang")
ids <- scheduleMlTasks(reg, files, iris)

showStatus(reg)
```

```{r eval=FALSE}
# retrieve result for first model once completed
m1 <- loadResult(reg, 1)
```

## Generator: R to XML

```{r}
set.seed(0)

data(iris)
predictors <- c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")

task <- newMlTask()

fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 3)

model <- task$train(x = iris[, predictors], y = iris$Species,
                    y_name = "Species",
                    method = "gbm",
                    preProcess = c("center", "scale"),
                    trControl = fitControl,
                    verbose = FALSE)

summary(model)

write(task, "iris_classification.xml")
unlink("iris_classification.xml")
```


## XML syntax of the Unified Machine Learning Language


```{r}
file_correct <- system.file("XML", "iris_classification.xml", package = "mlLang")
checkSchema(file_correct) # should work correctly

file_error <- system.file("XML", "iris_error.xml", package = "mlLang")
checkSchema(file_error) # should give a warning
```


## License

**mlLang** is released under the [MIT License](https://opensource.org/licenses/MIT)

Copyright (c) 2016 Andreas Frorath & Stefan Feuerriegel
